---
title: "Laporan Praktikum 6 AMP"
author: "Antonius Aditya Rizky Wijaya"
date: "2025-02-24"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Exercises

# Nomor 7

> In the chapter, we mentioned the use of correlation-based distance and Euclidean distance as dissimilarity measures for hierarchical clustering. It turns out that these two measures are almost equivalent: if each observation has been centered to have mean zero and standard deviation one, and if we let $r_{ij}$ denote the correlation between the $i$th and $j$th observations, then the quantity $1 - r_{ij}$ is proportional to the squared Euclidean distance between the ith and jth observations.
>
> On the `USArrests` data, show that this proportionality holds.

> _Hint: The Euclidean distance can be calculated using the `dist()` function, and correlations can be calculated using the `cor()` function._

```{r}
data("USArrests")
data_scaled <- t(scale(t(USArrests)))
euclidean_dist <- dist(data_scaled)^2
cor_dist <- as.dist(1 - cor(t(data_scaled)))
plot(euclidean_dist, cor_dist, 
     xlab = "Squared Euclidean Distance",
     ylab = "1 - Correlation",
     main = "Proportional Relationship")
```
Nilai (1 - korelasi Pearson) dan jarak Euclidean kuadrat menunjukkan pola yang mirip, ini membuktikan bahwa kedua ukuran dissimilarity ini hampir setara setelah standardisasi.


# Nomor 8

> In Section 12.2.3, a formula for calculating PVE was given in Equation 12.10. We also saw that the PVE can be obtained using the `sdev` output of the `prcomp()` function.
>
> On the `USArrests` data, calculate PVE in two ways:

> a. Using the `sdev` output of the `prcomp()` function, as was done in Section 12.2.3.

```{r}
data("USArrests")
hasil_pca <- prcomp(USArrests, center = TRUE, scale. = TRUE)
pve_a <- (hasil_pca$sdev^2) / sum(hasil_pca$sdev^2)
cat("PVE menggunakan sdev dari prcomp():\n")
print(pve_a)
```

> b. By applying Equation 12.10 directly. That is, use the `prcomp()` function to compute the principal component loadings. Then, use those loadings in Equation 12.10 to obtain the PVE. 
>
> These two approaches should give the same results.

```{r}
lambda <- apply(hasil_pca$x, 2, var)
pve_b <- lambda / sum(lambda)
cat("PVE menggunakan Equation 12.10:\n")
print(pve_b)

all.equal(pve_a, pve_b)
```
Menggunakan `sdev` dari `prcomp()`lebih sederhana daripada menghitung secara manual dengan Equation 12.10. Kedua metode memberikan hasil yang identik, yang membuktikan keakuratan teori PCA dalam menangkap variansi data.


### Question 9

> Consider the `USArrests` data. We will now perform hierarchical clustering on the states.
>
> a. Using hierarchical clustering with complete linkage and Euclidean distance, cluster the states.

```{r}
set.seed(42)
hc <- hclust(dist(USArrests), method = "complete")
```

> b. Cut the dendrogram at a height that results in three distinct clusters. Which states belong to which clusters?

```{r}
ct <- cutree(hc, 3)
sapply(1:3, function(i) names(ct)[ct == i])
```

> c. Hierarchically cluster the states using complete linkage and Euclidean distance, _after scaling the variables to have standard deviation one_.

```{r}
hc2 <- hclust(dist(scale(USArrests)), method = "complete")
```

> d. What effect does scaling the variables have on the hierarchical clustering obtained? In your opinion, should the variables be scaled before the inter-observation dissimilarities are computed? Provide a justification for your answer.

```{r}
ct <- cutree(hc, 3)
sapply(1:3, function(i) names(ct)[ct == i])
```

Scaling results in different clusters and the choice of whether to scale or 
not depends on the data in question. In this case, the variables are:

  - Murder    numeric  Murder arrests (per 100,000)  
  - Assault   numeric  Assault arrests (per 100,000) 
  - UrbanPop  numeric  Percent urban population      
  - Rape      numeric  Rape arrests (per 100,000)    

These variables are not naturally on the same unit and the units involved are
somewhat arbitrary (so for example, Murder could be measured per 1 million 
rather than per 100,000) so in this case I would argue the data should be 
scaled.

### Question 10

> In this problem, you will generate simulated data, and then perform PCA and $K$-means clustering on the data.
>
> a. Generate a simulated data set with 20 observations in each of three classes (i.e. 60 observations total), and 50 variables. 
>    
>    _Hint: There are a number of functions in `R` that you can use to generate data. One example is the `rnorm()` function; `runif()` is another option. Be sure to add a mean shift to the observations in each class so that there are three distinct classes._

```{r}
set.seed(42)
data <- matrix(rnorm(60 * 50), ncol = 50)
classes <- rep(c("A", "B", "C"), each = 20)
dimnames(data) <- list(classes, paste0("v", 1:50))
data[classes == "B", 1:10] <- data[classes == "B", 1:10] + 1.2
data[classes == "C", 5:30] <- data[classes == "C", 5:30] + 1
```

> b. Perform PCA on the 60 observations and plot the first two principal component score vectors. Use a different color to indicate the observations in each of the three classes. If the three classes appear separated in this plot, then continue on to part (c). If not, then return to part (a) and modify the simulation so that there is greater separation between the three classes. Do not continue to part (c) until the three classes show at least some separation in the first two principal component score vectors.

```{r}
pca <- prcomp(data)
ggplot(
  data.frame(Class = classes, PC1 = pca$x[, 1], PC2 = pca$x[, 2]),
  aes(x = PC1, y = PC2, col = Class)
) +
  geom_point()
```

> c. Perform $K$-means clustering of the observations with $K = 3$. How well do the clusters that you obtained in $K$-means clustering compare to the true class labels?
>    
>    _Hint: You can use the `table()` function in `R` to compare the true class labels to the class labels obtained by clustering. Be careful how you interpret the results: $K$-means clustering will arbitrarily number the clusters, so you cannot simply check whether the true class labels and clustering labels are the same._

```{r}
km <- kmeans(data, 3)$cluster
table(km, names(km))
```
$K$-means separates out the clusters nearly perfectly.

> d.  Perform $K$-means clustering with $K = 2$. Describe your results.

```{r}
km <- kmeans(data, 2)$cluster
table(km, names(km))
```
$K$-means effectively defines cluster 2 to be class B, but cluster 1 is a mix
of classes A and B. 

> e.  Now perform $K$-means clustering with $K = 4$, and describe your results.

```{r}
km <- kmeans(data, 4)$cluster
table(km, names(km))
```
$K$-means effectively defines cluster 1 to be class B, cluster 2 to be class A
but clusters 3 and 4 are split over class C.

> f.  Now perform $K$-means clustering with $K = 3$ on the first two principal component score vectors, rather than on the raw data. That is, perform $K$-means clustering on the $60 \times 2$ matrix of which the first column is the first principal component score vector, and the second column is the second principal component score vector. Comment on the results.

```{r}
km <- kmeans(pca$x[, 1:2], 3)$cluster
table(km, names(km))
```
$K$-means again separates out the clusters nearly perfectly.

> g.  Using the `scale()` function, perform $K$-means clustering with $K = 3$ on the data _after scaling each variable to have standard deviation one_. How do these results compare to those obtained in (b)? Explain.

```{r}
km <- kmeans(scale(data), 3)$cluster
table(km, names(km))
```
$K$-means appears to perform less well on the scaled data in this case.


# Nomor 11

> Write an `R` function to perform matrix completion as in Algorithm 12.1, and as outlined in Section 12.5.2. In each iteration, the function should keep track of the relative error, as well as the iteration count. Iterations should continue until the relative error is small enough or until some maximum number of iterations is reached (set a default value for this maximum number). Furthermore, there should be an option to print out the progress in each iteration.
> 
> Test your function on the `Boston` data. First, standardize the features to have mean zero and standard deviation one using the `scale()` function. Run an experiment where you randomly leave out an increasing (and nested) number of observations from 5% to 30%, in steps of 5%. Apply Algorithm 12.1 with $M = 1,2,...,8$. Display the approximation error as a function of the fraction of observations that are missing, and the value of $M$, averaged over 10 repetitions of the experiment.

```{r}

```


# Nomor 12

> In Section 12.5.2, Algorithm 12.1 was implemented using the `svd()` function. However, given the connection between the `svd()` function and the `prcomp()` function highlighted in the lab, we could have instead implemented the algorithm using `prcomp()`.
>
> Write a function to implement Algorithm 12.1 that makes use of `prcomp()` rather than `svd()`.

```{r}

```


# Nomor 13

> On the book website, `www.StatLearning.com`, there is a gene expression data set (`Ch12Ex13.csv`) that consists of 40 tissue samples with measurements on 1,000 genes. The first 20 samples are from healthy patients, while the second 20 are from a diseased group.
>
> a. Load in the data using `read.csv()`. You will need to select `header = F`.

```{r}
data <- read.csv("Ch12Ex13.csv", header = FALSE)
dim(data)
colnames(data) <- c(paste0("H", 1:20), paste0("D", 1:20))
```

> b. Apply hierarchical clustering to the samples using correlation-based distance, and plot the dendrogram. Do the genes separate the samples into the two groups? Do your results depend on the type of linkage used?

```{r}
hc.complete <- hclust(as.dist(1 - cor(data)), method = "complete") 
plot(hc.complete, main = "Dendrogram: Hierarchical Clustering (Complete Linkage)",
     xlab = "Sampel", ylab = "Jarak", sub = "", cex = 0.7)

hc.average <- hclust(as.dist(1 - cor(data)), method = "average")
plot(hc.average, main = "Dendrogram: Hierarchical Clustering (Average Linkage)",
     xlab = "Sampel", ylab = "Jarak", sub = "", cex = 0.7)

hc.single <- hclust(as.dist(1 - cor(data)), method = "single")
plot(hc.single, main = "Dendrogram: Hierarchical Clustering (Single Linkage)",
     xlab = "Sampel", ylab = "Jarak", sub = "", cex = 0.7)
```
Sampelnya terbagi menjadi dua kelompok, meskipun hasilnya agak bergantung pada metode hubungan yang digunakan. Complete linkage lebih sering menghasilkan grup yang lebih seimbang dibandingkan single linkage. Dalam kasus pengelompokan rata-rata, semua sampel sakit termasuk dalam satu subset dari sampel yang sehat. Jika hasil clustering tidak jelas, PCA mungkin diperlukan untuk mengurangi dimensi sebelum clustering.

> c. Your collaborator wants to know which genes differ the most across the two groups. Suggest a way to answer this question, and apply it here.

This is probably best achieved with a supervised approach. A simple method
would be to determine which genes show the most significant differences between
the groups by applying a t-test to each group. We can then select those with a
FDR adjusted p-value less than some given threshold (e.g. 0.05).

```{r}
class <- factor(rep(c("Healthy", "Diseased"), each = 20))
pvals <- p.adjust(apply(data, 1, function(v) t.test(v ~ class)$p.value))
which(pvals < 0.05)
```

